---
title: "Analysis of the National Park Visits"
subtitle: "Data Analytics 200 Team Project<br><br>" 
author: 
  - name: Gabrielle Swift
    affiliations:
      - name: SUNY Geneseo
  - name: Daniel West
    affiliations:
      - name: SUNY Geneseo
  - <br>
date: "`r Sys.Date()`"

format: 
  html:
    toc: true
    number-sections: true
    code-fold: false       # https://quarto.org/docs/output-formats/html-code.html
    code-summary: "Show the code"
    code-tools: true
    highlight-style: espresso # atom-one tango espresso

execute: 
  warning: false
  message: false
  echo: true
  eval: true
  fig-width: 9
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(scales)
library(kableExtra)
library(plotly)
library(viridis)


theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Introduction
National Parks are some of the best places for visitors and residents to explore nature and remember historical events. The United States is lucky to be home to several of these parks. The United States National Parks employ over 20,000 individuals. The amount visitors spend at these parks contributes to the economy. There are over 300 National Parks located throughout the United States, including in U.S. territories of the American Samoa, Puerto Rico and Guam. The National Park Service is primarily funded by Congress, however they also gain funds from park entrance fees and other private philanthropies. For years, the National park Service has lacked the funding it needs to properly protect these parks. The first National Park, created in 1872, was Yellowstone in Wyoming. These National Parks need sufficient funding in order to maintain the beauty of the parks and keep them from deteriorating so future generations can continue to enjoy these beautiful places.


My goal is to analyze the amount of visitation in these parks each year from 1979 to the most recent data, 2021. My analysis focuses on where these parks are located, by state and region and determine which parks contribute the most to the economy and which parks need to increase visitation. I focused on finding correlations between visitation and location as well as economic contribution and visitor spending. I also wanted to look at what impacts COVID-19 had on visitation.

I was able to figure out how visitation, park size, park type, and location correlated to the
amount of visitor spending and economic contribution. If I had more time, I would’ve liked to gather data on climate change, stocks, and financial patterns to see how these factors could have
affected visitation of parks.


# Data

I used data from the National Park Service from 1979 to 2021 to find correlations between visitors, states, parks and regions. I had to scrape data from the National Park Service webpage for these 30 years. I used Python and SQLite to glue the data sets together. Then I brought the data into Rstudio to do more analysis. This data includes visitation for each year and month. This was the primary data I used in my analysis. The Recreation Visits variable is the amount of visitors who enter the park, for example, this can be school groups, and tourists. Non Recreation Visits would refer to commuters traveling through parks, and NPS employees. I focused on the recreation visits, to show a more realistic count of the visitation at each park. The Region variable refers to the subdivisions of the United States as administered by the National Park Services. It includes 7 different regions including Alaska, Intermountain, Midwest, National Capital, Northeast, Pacific West, Southeast and Northeast. The National Capital is part of the Northeast but it is a separate region, only including Washington DC. Another interesting variable was Park Type, which included 17 different park types, ranging from National Battlefield, to National Monuments, Parks, and Seashores. Most of these variables were character variables so I cleaned the data by transforming Year and Month variables into integers, and Recreation Visits as numeric. This was a very large dataset, for 30 years of data, I had 173,560 observations and 30 variables. When I started summarizing the data I took a look at the number of parks located in the United States from 1979 to 2021.


I also used All Trails data to see which parks had the most trails and the exact coordinates of these trails and parks. I obtained the All Trails data from Kaggle. All Trails is a website, created for hikers. It acts as a platform to help find the best hiking, running, or biking trails and includes information like reviews, ratings, activities offered, and coordinates of trails. For the All Trails data, I was able to separate latitude and longitude into separate variables in Python so I could use them to make a map visualization. The All Trails data does not include every state, only ones with trails. If I was given more time, I would’ve liked to look more into the reasons why certain parks are not listed in the dataset. My hypothesis is either the data is not up to date or they only list trails that are listed as hiking trails, rather than trails and walkways that can be found at historic sites, and monuments. While performing summary statistics I was able to find which state had the most trails.


I also included Visitor Spending from 2020 to see how visitation, visitor spending and economic output vary by each state, during COVID-19. The National Park Service provided visitor spending effects, by national economy, state economy, and park economy. However, the Park Data was not readily available to put into R, so I used the visitor spending per state. This data is important in the economic analysis of my data. It shows me which states have the most amount of visitors, how much revenue they got through visitor spending, and how much each state contributed to the economy. This information can also be used to see how the parks help drive tourism and increase outdoor recreation activities. Visitor spending at the parks helps support jobs and benefits the nation's economy. I would have liked to gather more information on the jobs and visitor spending for the 30 year period I looked at to see how it has changed over the years, and if the economic contribution is impacted by any other factors. Given more time I would have also looked at which parks accumulated the most revenue each year.


## Motivation For The Project 

I want to look at National Park data for my research project. The National Parks are an important part of the United States and function as areas of land that are preserved by the national government. They can have historical elements, or be a place to enjoy the wonderful views they have to offer. I have been to several National Parks myself. I would like to look at the data to : 

  - Find which parks have the most visitors each year and if visitation in the parks is increasing or decreasing. 
  
  - Identify what could have caused these spikes and peaks in visitation.
  
  - Explore which regions of the United States have the most parks. 
  
  - Determine if certain parks contribute more to the economy. 
  
  - Look at the impacts COVID-19 had on visitation. 

## Background Knowledge
The National Park Service is primarily funded by Congress, but also is funded through park entrance fees and some private philanthropies. For years the National Park Service has lacked the funding it needs to maintain and protect these national parks. Since the first National Park, Yellowstone, was established in 1872, there have been an additional 400 National Parks with over 20,000 employees added. 

This research is important, because it can help determine how much funding the National Parks need and which ones need more attention than others. It can help us discover how different issues have affected these National Parks. It can also help us see if certain parks should be closed during certain times of the year in order to have a more efficient and effective budget for these parks. 

We know that national parks provide peaceful places to enjoy scenery and give wildlife and native plants a safe home which maintains our ecosystem. Economically, national parks create jobs in tourism, park management and capital works and draw visitors to regional areas where they spend their money in local towns. 

## Summary of the Modeling Result 
In my model I was able to figure out how visitation, park size, and park type correlated to the amount of visitor spending and economic contribution of the parks. If I had more time, I would've liked to gather data on climate change, stocks, and financial patterns to see how these factors could have affected visitation at parks. I was able to see that visitation decreased in 2020 due to the COVID-19 pandemic. 

# Data Summary 
I was able to collect 30 years of data from the National Park Service. I collected **Public Use Data** data from 1979 to 2021. I also collected data on the amount of trails located in each park from Kaggle, **All Trails Data**. Additionally I was able to find **Visitor Spending Data** for 2020. 

## All Trails Data  
  - This data is from the All Trails website, which is a platform containing all hiking trails located in each National Park. It includes the coordinates of each trail, park name, state name, and city name. Additionally this data provides variables like the average rating, length, elevation gain, and difficulty rating for each trail. 
  
  
### Summary Statistics: All Trails

```{r, message = FALSE}
#Read csv file 
alltrails <- read_csv("/Users/bchoe/My Drive/suny-geneseo/teaching-materials/lecture-data/national-park-visits.csv")

Recreation.Visits <- alltrails$`Recreation.Visits`

#first 6 rows 
head(alltrails)

rmarkdown::paged_table(alltrails)
```
```{r}
#Number of Trails per state
TrailsperPark <-alltrails%>%
  group_by(State)%>%
  summarize(trails = n_distinct(Park))
TrailsperPark%>%
  arrange(desc(trails))
```

```{r}
nvars <- format(round(ncol(alltrails), 0), nsmall = 0, big.mark = ",")
nobs <- format(round(nrow(alltrails), 0), nsmall = 0, big.mark = ",")
```

The number of variables is `r nvars`; the number of observations is `r nobs`. 

## National Park Service Data 


### Public Use Statistics (1979-2021)
  - Public Use Statistics give the number of visits per month and year for each National Park. This data includes Recreation and Non-Recreation Visits, as well as the hours the park was open for each group and the number of visitors who camped, stayed overnight, etc. 

```{r}
skim(alltrails) %>% 
  select(-n_missing)

```

```{r}
#Total Visits per Year in each Region 
alltrails%>%
  group_by(Year, Region)%>%
  summarize(Total_Visits = n_distinct(Recreation.Visits))
```

## Economic Contribution of National Park Visitor Spending 

### Visitor Spending Data 
  - Gives us information on the number of visitors in each park and how much they spend in each National Park, the number of jobs provided by parks, and home much income workers make. 

```{r}
Visitor_Spending_2022 <- read_csv("/Users/bchoe/My Drive/suny-geneseo/teaching-materials/lecture-data/national-park-spending.csv")
head(Visitor_Spending_2022)
skim(Visitor_Spending_2022) %>% 
  select(-n_missing)
```

# Research Questions

  - Is visitation increasing or decreasing at certain parks? 
  - How many parks are in the United States? Each Region? Each State? 
    - How many trails are in each park? 
  - Are there seasonal trends? 
  - Does the location of the park correlate to the amount of visitation? 
  - Does the size and amount of parks correlate with the economic contribution of the park? 


# Hypothesis 

I believe that parks located in the West will get more visitation and result in more economic contribution than parks located in the East. 

  - The West has more National Parks because at the time when the National Park System was established, most of the land in the East was privately owned.
  
  - The West is less urbanized and has more open land.
  - I think California will have the most parks. 
  

## How many Parks are in the United States? 
```{r}
total_park_count <-alltrails%>%
  group_by(Year)%>%
  summarize(p = n_distinct(Park))
uspark <- ggplot(data = total_park_count, 
       mapping = aes(x = Year, y = p))+ 
  geom_line(color="#69b3a2", size=2)+ 
  ylim(100, 400)+
  ggtitle("Number of National Parks in the United States \n(1979 - 2021)")+
  xlab("Year")+
  ylab("Total Number of Parks")
uspark
```
The number of Parks in the United States has increased from 268 in 1979 to 368 in 2021. 


## Number of Parks Per Region
```{r}
numparks_region<-alltrails%>%
  group_by(Region)%>%
  summarize(Parks = n_distinct(Park))

numparks_region%>%arrange(desc(Parks))

```
The number of parks found in each region. 

To see a map of the Regions click [here](Region Map.jpeg.pdf).

## Visits By Region 
```{r, message = FALSE}
regionvis <- alltrails%>%
  group_by(Region, Year)%>%
  summarize(visits = sum(Recreation.Visits))
regionvis
ggplot(data = regionvis, 
       mapping = aes(x = Year, y =visits, fill = Region))+
  geom_col()

faceted_region <- ggplot(data = alltrails, 
                         mapping = aes(x = Year, y = Recreation.Visits))+
  geom_point(alpha = 0.2)+
  facet_wrap(~Region, ncol = 3)
faceted_region+
  theme_bw()+ 
  ggtitle("Recreation Visits By Region")+ 
  ylab("Recreation Visits")+ 
  xlab("")+
  scale_y_continuous(labels = comma)
```


## Number of Parks per State and Region
```{r, warning = FALSE, message=FALSE}
park_state <- alltrails%>%
  group_by(State, Region)%>%
  summarize(park = n_distinct(Park))
park_state%>% arrange(desc(park))%>%ggplot(data = park_state, 
       mapping = aes(x = State, y = park, color = Region))+ 
  geom_point()+ 
  theme(axis.text.x = element_text(angle = 90))+ 
  labs(x = "State", y = "Park Count", 
       title = "Amount of Parks by State and Region")
```

## Number of Trails Per Park 
```{r}
TrailsperPark <-alltrails%>%
  group_by(State)%>%
  summarize(trails = n_distinct(Park))%>%
  arrange(desc(trails))
c <- ggplot(TrailsperPark, aes(x = State, 
                               y = trails))+
  geom_segment(aes(x = State, xend = State, y = 0, yend = trails),
               color = "skyblue") +
  geom_point(color = "blue", size = 4, alpha = 0.6)+
  theme_light()+
  coord_flip()+
  theme(panel.border = element_blank(),
        axis.ticks.y = element_blank())+
  ggtitle("Number of Trails in each State")+
  xlab("")+
  ylab("Number of Trails")
c
```




## Visitation in 2021
```{r}
lastyear <- alltrails%>%
  filter(Year == 2021)%>%
  select(Year, Month, Park, Region, Recreation.Visits)
ggplot(data = lastyear, 
       mapping = aes(x = Month, y = Recreation.Visits, fill = Region))+
  geom_bar(position = "stack", stat = "identity")+
  scale_x_continuous(breaks = breaks_width(1))+
  theme_classic()+
  scale_y_continuous(labels = label_number(suffix = "K", scale = 1e-6))+
  ylab("Recreation Visit (Thousands) ")+ 
  xlab("Month")+
  ggtitle("Visitation By Month in 2021")
```




# Conclusion 
The main findings were that California has the most amount of parks and trails, though it is not located in the region that has the most parks. California also has the most amount of visitors per year, so this correlates to why they have the most amount of visitor spending and the highest economic contribution. We can conclude that California does not have to rely heavily on government funding. States with the most parks tend to have higher rates of visitation, and therefore the most visitor spending and economic contribution. I would like to look further into why Washington DC has less economic contribution and visitor spending. My thoughts are that a lot of visits are school bus trips, thus the students are not spending as much as a family would on their visits to the parks. If parks are able to increase their visitation, then they would be able to contribute more to the economy and generate more revenue that could in turn increase the funding that goes toward maintenance of these parks. We saw that the COVID-19 pandemic affect the parks, by decreasing visitation in all regions. I would like to look further to see if any specific parks increased in visitation during these times.


I had many limitations while conducting my research, such as not gathering all the data I wanted and having trouble scrapping the specific data. Gathering and cleaning the data took a lot longer than anticipated and therefore set me back in my research and development of economic analysis. I hope to continue this research and look further into variables that made me question the outcomes. I would like to see how visitation changed for each park over the years, and in each month to see if certain parks and park types have become more or less popular over the years. I would also like to look into the climate changes, and travel patterns to see if these factors have to do with visitation rates. I would look further into the average rating of parks and traffic counts in parks. If I can gather all the variables I need, my analysis would be more thorough and be able to make better suggestions to how the parks can change their current mannerisms to make the National Park System better. For now, we can continue to visit these parks and contribute to the maintenance needed. We can do our part by cleaning up parks while we enjoy hikes, to conserve the beauty displayed at these parks.

